---
title: "ImPRESS Statistical Analysis Mock-up report Report"
author: "ImPRESS Statistical Team"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: '2'
  word_document:
    toc: true
    toc_depth: 2
  html_document:
    toc: true
    toc_depth: 2
subtitle: Baseline Review for the Newly Diagnosed Glioblastoma Cohort
---

# Introduction

The ImPRESS study is a prospective investigation of losartan as an adjunctive therapy for patients with glioblastoma. According to Protocol Version 7.0 (17-Dec-2024), the trial collects comprehensive demographic, clinical, imaging, and quality-of-life information prior to treatment initiation. The statistical objectives for the interim analyses focus on characterising baseline status, neurological function, and tumour burden before randomisation to evaluate comparability across planned treatment arms.

This report summarises the currently available baseline data with emphasis on clinical status, imaging assessments, and prior treatments that may influence subsequent outcomes. The material is intended to provide early insight into data quality and cohort composition ahead of formal efficacy or safety analyses.

# Randomisation Status

The present dataset is built on *sham randomisation* generated from the most recent export (see `_targets.R` workflow). No true treatment assignments have been unlocked. All findings in this document should therefore be interpreted as preliminary and descriptive only. Final efficacy comparisons will not be conducted until the database is locked and genuine randomisation codes are released.

# Cohort Scope

The results herein are restricted to participants in cohort code `r cfg$analysis$cohortcd` as configured in `config/cfg.yml`. Subjects from other cohorts are excluded from all summaries, figures, and tables in this draft. A combined-cohort analysis can be delivered once recruitment and cleaning are complete for all strata.

# Baseline Characteristics

This section summarises baseline characteristics for the newly diagnosed glioblastoma cohort using the `adbl` ADaM dataset.

```{r setup, message=FALSE, warning=FALSE, echo = FALSE, include = FALSE}
library(targets)
library(dplyr)
library(tidyr)
library(purrr)
library(rlang)
library(flextable)
library(rtables.officer)
library(rtables)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)
options(tidyverse.quiet = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())



```

```{r load-data, message=FALSE, warning=FALSE}
tar_load(cfg)
tar_load(tbl_baseline)
tar_load(primary_mri_sections)
tar_load(primary_mri_labels)
tar_load(other_mri_summaries)
tar_load(cycle11_mri_summaries)
tar_load(neuro_summaries)
tar_load(qol_summaries)
tar_load(steroid_summaries)
tar_load(rano_summaries)
tar_load(tte_sections)
tar_load(lb_summaries)
tar_load(lb_late_summaries)
tbl_baseline |> 
  tt_to_flextable() %>%
  autofit()

```

# Primary Efficacy Analyses (MRI)

The primary variable is analysed with the MCP-Mod methodology as described in the protocol and SAP. We pre-specify four candidate dose-response models (linear, Emax, sigmoid Emax, logistic) and derive optimal contrasts using the estimated variance-covariance from the mixed model. The MCP step tests for a dose-response signal across candidate models with multiplicity control; if significant, the Mod step selects the best-fitting model (by AIC) and reports ED values. Results below are grouped by MRI endpoint (primary, key secondary, sensitivity). All analyses are based on data up to and including Day 43 (Week 6).

## Candidate dose-response models

```{r mri-models, message=FALSE, warning=FALSE}
print(primary_mri_sections[[1]]$plots$candidate_models)
```


```{r mri-primary, message=FALSE, warning=FALSE, results='asis'}


for (nm in names(primary_mri_sections)) {
  sec <- primary_mri_sections[[nm]]
  lbl <- primary_mri_labels[[nm]] %||% nm

  cat("## ", lbl, "\n\n", sep = "")

  cat(sec$text$intro, "\n\n")
  cat(sec$text$nsubj, "\n\n")
  cat(sec$text$models, "\n\n")


  cat("### Line plot \n\n")
  print(sec$plots$lineplot)
  cat("\n\n") 

  cat("### Optimal contrasts\n\n")
  print(sec$tables$contmat)
  cat("\n\n")

  cat("### MCP-Mod tests\n\n")
  print(sec$tables$mct)
  cat("\n\n")

  cat("### Fitted model parameters (AIC/ED)\n\n")
  print(sec$tables$fit)
  cat("\n\n")

  cat("### Estimated means by dose\n\n")
  print(sec$tables$estimates)
  cat("\n\n")

  cat("### Fitted dose-response curves\n\n")
  purrr::walk(sec$plots$fit_models, print)
  cat("\n\n")

  if (!is.null(sec$tables$anydose)) {
    cat("### Supplementary analysis: any dose vs 0 mg\n\n")
    print(sec$tables$anydose)
    cat("\n\n")
  }

}
```

# Other MRI Endpoints up to Day 43

For remaining MRI endpoints, we present estimated means by dose, contrasts vs 0 mg, and a plot with confidence intervals. 

```{r mri-other, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(other_mri_summaries)) {
  obj <- other_mri_summaries[[nm]]
  if (is.null(obj)) next
  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " subjects were included in the analysis for this endpoint.\n\n", sep = "")
  cat("### Line plot by treatment start\n\n")
  print(obj$lineplot)
  cat("\n\n") 
  cat("### Means by dose\n\n")
  print(obj$means)
  cat("\n\n### Differences vs 0 mg\n\n")
  print(obj$diffs)
  cat("\n\n### Plot\n\n")
  print(obj$plot)
  if (!is.null(obj$anydose)) {
    cat("\n\n### Supplementary analysis: any dose vs 0 mg\n\n")
    print(obj$anydose)
  }
  cat("\n\n\\newpage\n\n")
}
```

# MRI Endpoints at Cycle 11 (Day 141)

For all MRI endpoints, we report estimated change from baseline at Cycle 11 (day 141), grouped by the extended treatment variable. The summaries include mean estimates, contrasts versus the reference group, and a plot with confidence intervals.

```{r mri-cycle11, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(cycle11_mri_summaries)) {
  obj <- cycle11_mri_summaries[[nm]]
  if (is.null(obj)) next
  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " subjects were included in the analysis for this endpoint.\n\n", sep = "")
  cat("### Means by treatment\n\n")
  print(obj$means)
  cat("\n\n### Differences vs reference\n\n")
  print(obj$diffs)
  cat("\n\n### Plot\n\n")
  print(obj$plot)
  cat("\n\n\\newpage\n\n")
}
```

# Neurologic Performance

Change from baseline in KPS, ECOG, and NANO scores at Days 15/43 and model-based estimates at Days 155/239 are reported below for the newly diagnosed glioblastoma cohort. Note that since the day 29 is not measured for any of the measures, the dose-response analyses are based on Days 15 and 43 only for the early timepoints. There might therefore be a limited ability to detect dose-response relationships at these early visits becasue we are lacking one contrast. 



```{r neuro-performance, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(neuro_summaries)) {
  obj <- neuro_summaries[[nm]]
  if (is.null(obj)) next
  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " subjects were included in the analysis for this endpoint.\n\n", sep = "")

  if (!is.null(obj$early)) {
     if (!is.null(obj$early$plot_step)) {
      cat("\n\n### Line plot by treatment start\n\n")
      print(obj$early$plot_step)
    }
    cat("\n\n### Days 15/43 (change from baseline)\n\n")
    print(obj$early$means)
    cat("\n\n### Differences vs 0 mg\n\n")
    print(obj$early$diffs)
    cat("\n\n### Plot\n\n")
    print(obj$early$plot)
   
    cat("\n\n")
  }

  if (!is.null(obj$late)) {
    cat("### Days 155/239 (mixed model)\n\n")
    print(obj$late$means)
    cat("\n\n### Differences vs reference\n\n")
    print(obj$late$diffs)
    cat("\n\n### Plot\n\n")
    print(obj$late$plot)
    cat("\n\n")
  }

  cat("\n\n\\newpage\n\n")
}
```

# Quality of Life Outcomes

Model based estimates of change from baseline in QLQ-C30 and QLQ-BN20 scores at Days 155/239 are reported below for the newly diagnosed glioblastoma cohort. Note that even if the QoL outcomes were collected at Day 43 it is not possible to make a meaningful analysis of the effect of losartan as all participants received the treatment at that time point.

```{r qol-outcomes, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(qol_summaries)) {
  obj <- qol_summaries[[nm]]
  if (is.null(obj)) next
  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " subjects were included in the analysis for this endpoint.\n\n", sep = "")

  if (!is.null(obj$early)) {
    cat("### Days 15/29/43 (change from baseline)\n\n")
    print(obj$early$means)
    cat("\n\n### Differences vs 0 mg\n\n")
    print(obj$early$diffs)
    cat("\n\n### Plot\n\n")
    print(obj$early$plot)
    cat("\n\n")
  }

  if (!is.null(obj$late)) {
    cat("### Days 155/239 (mixed model)\n\n")
    print(obj$late$means)
    cat("\n\n### Differences vs reference\n\n")
    print(obj$late$diffs)
    cat("\n\n### Plot\n\n")
    print(obj$late$plot)
    cat("\n\n")
  }

  cat("\n\n\\newpage\n\n")
}
```

# Steroid Use

Binary steroid use is analysed with longitudinal logistic regression and population-averaged marginal risk differences for early (Days 15/29/43) and late (Days 155/239) visits in the newly diagnosed glioblastoma cohort.

```{r steroid-use, message=FALSE, warning=FALSE, results='asis'}
obj <- steroid_summaries
if (!is.null(obj)) {
  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " subjects were included in the analysis.\n\n", sep = "")

  if (!is.null(obj$early)) {
    cat("### Days 15/29/43 (by dose)\n\n")
    print(obj$early$means)
    cat("\n\n### Risk differences vs 0 mg\n\n")
    print(obj$early$diffs)
    cat("\n\n### Plot\n\n")
    print(obj$early$plot)
    cat("\n\n")
  }

  if (!is.null(obj$late)) {
    cat("### Days 155/239 (population-averaged risk)\n\n")
    print(obj$late$means)
    cat("\n\n### Risk differences vs reference\n\n")
    print(obj$late$diffs)
    cat("\n\n### Plot\n\n")
    print(obj$late$plot)
    cat("\n\n")
  }

  cat("\n\n\\newpage\n\n")
}
```

# RANO Responses

Ordinal RANO outcomes are analysed using proportional odds models within the configured cohort. For `r rano_summaries$trsdisea$label`, visits at Days 141 and 225 are included; for `r rano_summaries$trorresp$label`, only Day 225 is included. Proportional odds ratios and confidence intervals are reported below.

```{r rano-responses, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(rano_summaries)) {
  obj <- rano_summaries[[nm]]
  if (is.null(obj)) next
  cat("## ", obj$label, "\n\n", sep = "")
  for (tbl in obj$tables) {
    cat("### Visit ", tbl$avisitn, "\n\n", sep = "")
    print(tbl$table)
    cat("\n\n")
  }
  cat("\n\n\\newpage\n\n")
}
```

# Biomarker Endpoints (Serum Cytokines / Angiogenic Factors)

Serum cytokine and angiogenic-factor concentrations from Bio-Plex 19-plex and 14-plex
panels are analysed as exploratory biomarker endpoints.  For each analyte, a linear
mixed model with random subject intercept is fitted with change from Screening baseline
as the response.  Where the log-scale model fits better (lower AIC), the analysis is
performed on the log-concentration scale and results are labelled accordingly.

Biomarker samples are collected at screening (baseline), Day 15 (Cycle 2), and Day 43 (Cycle 4) for the early analysis, and again at Cycle 17 (Day 225) and later for the long-term analysis. The early analysis uses the stepped-wedge treatment assignment; the long-term analysis uses the assigned long-term dose (`trt02p`).


A formal model selection and residual diagnostic assessment is available in
a separate document (impress_lb_model_fit.pdf).

```{r lb-outcomes, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(lb_summaries)) {
  obj <- lb_summaries[[nm]]
  if (is.null(obj)) next

  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " participants contributed data for this analyte.\n\n",
      sep = "")

  cat("### Estimated change from baseline by dose\n\n")
  print(obj$means)
  cat("\n\n### Differences vs 0 mg\n\n")
  print(obj$diffs)
  cat("\n\n### Model comparison (AIC)\n\n")
  print(obj$aic)
  cat("\n\n### Plot\n\n")
  print(obj$plot)

  cat("\n\n\\newpage\n\n")
}
```

# Biomarker Endpoints — Long-term (Cycle 15+)

Model-based estimates of change from baseline at Cycle 15 and later are reported below for analytes with available long-term samples. The same scale (linear or log-transformed) as selected for the early analysis is used for each analyte.

```{r lb-late-outcomes, message=FALSE, warning=FALSE, results='asis'}
for (nm in names(lb_late_summaries)) {
  obj <- lb_late_summaries[[nm]]
  if (is.null(obj)) next

  cat("## ", obj$label, "\n\n", sep = "")
  cat("A total of ", obj$nsubj, " participants contributed data for this analyte.\n\n",
      sep = "")

  cat("### Estimated change from baseline by long-term dose\n\n")
  print(obj$means)
  cat("\n\n### Differences vs 0 mg\n\n")
  print(obj$diffs)
  cat("\n\n### Model comparison (AIC)\n\n")
  print(obj$aic)
  cat("\n\n### Plot\n\n")
  print(obj$plot)

  cat("\n\n\\newpage\n\n")
}
```

# Time-to-Event Endpoints

Kaplan-Meier curves are shown with the at-risk table below. Overall log-rank p-values are reported for each endpoint, followed by Cox regression hazard ratios versus 0 mg.

```{r tte-endpoints, message=FALSE, warning=FALSE, results='asis', fig.height=8}
for (nm in names(tte_sections)) {
  obj <- tte_sections[[nm]]
  if (is.null(obj)) next
  cat("## ", obj$label, "\n\n", sep = "")
  print(obj$km_plot)
  cat("\n\n### Log-rank p-value\n\n")
  cat("Overall log-rank p-value: ", obj$logrank_p, "\n\n", sep = "")
  cat("### Cox regression (vs 0 mg)\n\n")
  print(obj$cox_table)
  cat("\n\n\\newpage\n\n")
}
```
