---
title: "Dilafor PPL07 Statistical Report  for the Primary Analysis"
author: "Inge Christoffer Olsen"
date: "23 October 2018"
output:
  word_document: 
    toc: true
  html_document:
    number_sections: yes
    toc: true
  pdf_document:
    number_sections: yes
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
# Introduction
## Preface

This document presents the report of the findings from the primary efficacy analyses of the trial titled "Randomized, Double-Blind, Placebo-Controlled, Parallel-Group, Dose-Finding Study to Evaluate the Effect of continuous infusion of Tafoxiparin as an Adjunct Treatment to Oxytocin for up to 36  hours in Term Pregnant, Nulliparous Women to treat Primary Slow Progress of Labor including prolonged latent phase and Primary Labor Arrest". The analyses are performed as described in the protocol version 6.0, dated January 24, 2018, final signature approval date February 02, 2018, and detailed in the statistical analysis plan (SAP) version 1.0 dated 03Oct2018.

The report presents all output to be run on the primary endpoint. 

The primary efficacy analyses consists of assessing the dose-response relationship of the primary endpoint. Other secondary, exploratory and safety analyses will be reported elsewhere.

# Objectives and endpoints
## Primary objective
The primary objective of this study is to assess the dose-response relationship of tafoxiparin as an adjunct to oxytocin therapy on the labor time defined as the time from the start of continuous infusion of tafoxiparin/placebo, until partus in term-pregnant, nulliparous women requiring labor augmentation due to Primary Slow Progress of Labor including prolonged latent phase or Primary Labor Arrest.

## Primary endpoint
The primary endpoint is the time from start of continuous infusion of tafoxiparin/placebo until vaginal partus.

# Analysis datasets
There are two analysis datasets used for the primary analyses:

* Full Analysis Set (FAS) – This is the primary population and comprises of all eligible and randomized study patients having received any study treatment. Efficacy analyses will be performed against this population. The FAS will be based on randomized treatment allocation and not by actual treatment given (if different), in accordance with the intention to treat (ITT) principle.
* Per Protocol (PP) Population - This will include all patients in the study who report no major protocol violations affecting the efficacy evaluation.

The allocation of each included patient to either of the analysis datasets was done during the blind review of the data.


```{r import, include = FALSE}
library(haven)
adef <- read_sas("5.3 Analyser/Datasets/raw/TO_INGE_19OCT2018/adef2.sas7bdat", NULL)
adef1 <- read_sas("5.3 Analyser/Datasets/raw/TO_INGE_19OCT2018/adef1.sas7bdat", NULL)

library(tidyverse)
# Filter, clean and select variables

adprim <- adef %>%
  rename_all(tolower) %>%
  filter(parcat1n==1) %>% #select primary variable
  filter(fasfl=="Y") %>% # Select FAS
  modify_at(c("siteid", "sitegrpc", "trtp", "indcdc"),as.factor) %>%
  mutate(trtp = fct_reorder(trtp,trtpn), 
         sitegrprc = fct_reorder(sitegrpc,sitegrpn)) %>%
  select(usubjid, siteid, sitegrpc, sitegrpn, trtp, trtpn, aval, 
         cnsr, indcdc, indcdn, fasfl, pprotfl)

```


There were `r sum(adprim$fasfl=="Y")` patients included in the Full Analysis Set, and `r sum(adprim$pprotfl == "Y")` patients included in the Per Protocol Population. The distribution by treatment is given in Table 1. Details of exclusions from the analysis datasets will be provided elsewhere.

```{r alloc, echo=FALSE}

f_alloc <- function(data){
data %>%
  group_by(trtp) %>%
  summarise(nfas=sum(fasfl=="Y"), 
            npps=sum(pprotfl=="Y")) %>%
  ungroup() %>%
  mutate(txtfas=paste(nfas," (", round(nfas/sum(nfas)*100,digits=1),")")) %>%
  mutate(txtpps=paste(npps," (", round(npps/sum(npps)*100,digits=1),")")) %>%
  add_row(trtp="Total", .before=1) %>%
  mutate(txtfas = paste(ifelse(trtp == "Total", sum(nfas,na.rm=TRUE), txtfas))) %>%
  mutate(txtpps = paste(ifelse(trtp == "Total", sum(npps,na.rm=TRUE), txtpps))) %>%
  select(trtp, txtfas, txtpps) %>%
  knitr::kable( col.names=c("Treatment", "FAS, n (%)", "PPS n (%)"), caption = "Table 1")
  
}
prim <- list()
prim$alloc <-  f_alloc(adprim)
prim$alloc

```

# Descriptive statistics of the primary variable
Descriptive analysis of the primary variable includes Kaplan-Meier plots and corresponding quantiles by treatment group.

```{r desc, echo=FALSE, message=FALSE}
library(survival)
library(survminer)
library(ggplot2)

f_km = function(x) {

  sfit <- survfit(Surv(aval,cnsr == 0) ~ trtp, data=x)
  kmplot <- ggsurvplot(sfit,
    data=x,
    size=0.5,
    fun="event",
    palette="lancet",
    risk.table=TRUE,
    risk.table.height = 0.35,
    legend=c(0.8,0.4),
    legend.labs = c("Placebo", "5 mg/h", "15 mg/h", "25 mg/h"),
    xlab="Time (hours)",
    ylab="Vaginal delivery probability",
    pval=FALSE,
    legend.title = ""
  )
  quantile <- sfit %>%
    quantile(probs=c(0.25,0.5,0.75)) %>%
    .$quantile %>%
    t %>%
    as_tibble(rownames=NA) %>%
    rownames_to_column %>%
    mutate(rowname = c("Q1", "Median", "Q3")) %>%
    knitr::kable(
      col.names = c("", "Placebo", "5 mg/h", "15 mg/h", "25 mg/h"),
      digits=2,
      caption="Quantiles from the Kaplan-Meier plot"
    )

  list(sfit=sfit,kmplot=kmplot,quantile=quantile)
}

prim$desc <- f_km(adprim)
prim$desc$kmplot
prim$desc$quantile
```


# Primary analysis of efficacy
The primary analysis follows the MCP-Mod methodology as detailed in the SAP. The procedure consists of five steps to evaluate the dose-response relationship.

## Step 1: The candidate dose response models
The four pre-specified candidate dose-response models are detailed in the SAP, and are 1) a linear model; 2) an E~max~ model with ED~50~ = 12 mg/h; 3) a sigmoid E~max~ model with ED~50~ = 12 mg/h and Hill = 5 and 4) a logistic model with E~50~ = 15 mg/h and $\delta$ = 3.33 mg/h. The candidate models are depicted in Figure 1.

```{r drplots, echo=FALSE, message=FALSE}
library(DoseFinding)
doses <- c(0,5,15,25)
fmodels <- Mods(linear = NULL, emax = c(12), sigEmax=c(12,5), logistic=c(15,3.33),
                doses = doses, placEff=9, maxEff=-9*0.25)
fmodels_names <- names(fmodels)
names(fmodels_names) <- names(fmodels)
plot(fmodels, ylab = "Median time to delivery", main="Figure 1: Pre-specified candidate dose-response models")
```

## Step 2: Optimal contrasts from the candidate models
The optimal contrasts are derived from the candidate response models with the pre-specified model parameters from step 1, in addition to the estimated variance-covariance matrix $\boldsymbol S$ of the estimated dose response parameters (median time to vaginal partus) from the observed data.


```{r primest, echo=FALSE}
library(survival)
library(msm)
library(broom)

f_est <- function(a_data, a_sreg) {

  sreg <- a_sreg
  mean <- coef(sreg)
  mean <- mean[str_detect(names(mean),"^trtp")]
  term <- str_replace(names(mean),"^trtp","")
  median <- exp(mean)*(log(2)^sreg$scale)
  trtnames <- names(mean)
  S <- vcov(sreg)[trtnames,trtnames]
  S_median <- deltamethod(
    g= list(
      ~exp(x1)*(log(2)^exp(x5)),
      ~exp(x2)*(log(2)^exp(x5)),
      ~exp(x3)*(log(2)^exp(x5)),
      ~exp(x4)*(log(2)^exp(x5))
      ),
    mean=c(mean,log(sreg$scale)),
    cov=vcov(sreg)[c(trtnames,"Log(scale)"),c(trtnames,"Log(scale)")],
    ses=FALSE
    )

  table <- tidy(sreg) %>%
    filter(str_detect(term,"^trtp")) %>%
    mutate(median = exp(estimate)*(log(2)^sreg$scale)) %>%
    mutate(term = str_replace(term,"^trtp","")) %>%
    mutate(se_median = sqrt(diag(S_median))) %>%
    mutate(median_txt = paste0(
      round(median,digits=2),
      " (",
      round(median - qnorm(0.975)*se_median, digits=2),
      " to ",
      round(median + qnorm(0.975)*se_median, digits=2),
      ")"
      )) %>%
    mutate(mean_txt = paste0(
      round(estimate, digits=2),
      " (",
      round(conf.low, digits=2),
      " to ",
      round(conf.high, digits=2),
      ")"
      ))
  
  list(mean = mean,
       median = median,
       S = S,
       S_median = S_median,
       table = table,
       term = term)
}


prim$sreg <-survreg( Surv(aval, cnsr == 0) ~ 
                       trtp -1 + indcdc + sitegrpc, data = adprim, dist="weibull")

prim$est <- f_est(adprim, prim$sreg)

```

The transformed estimates of median time to delivery from the Weibull scale parameters and the variance-covariance matrix are given by

```{r prim_estres, echo=FALSE}
prim$est$table %>%
  select(term,median_txt) %>%
  knitr::kable( col.names=c("Treatment" ,"Median time to delivery (95% CI)"), caption="Transformed estimates")

prim$est$S_median %>%
  as_tibble() %>%
  rownames_to_column(var="S") %>%
  mutate(S = prim$est$term) %>%
  rename_at(vars(V1:V4), funs(prim$est$term)) %>%
  knitr::kable(digits=3, caption="Variance-covariance matrix")
```


Together with the candidate response models, the optimal contrasts are given by
```{r prim_optcont, echo=FALSE}
library(DoseFinding)
f_contmat <- function(dfobj) {

  contMat <- optContr(fmodels, S=dfobj$est$S_median)
  table <-contMat$contMat %>%
  as_tibble(rownames=NA) %>%
  rownames_to_column(var="Treatment") %>%
  mutate(Treatment = dfobj$est$term) %>%
  knitr::kable(digits=3)

  list(contmat = contMat, table = table)
}

prim$contmat <- f_contmat(prim)
prim$contmat$table

```

## Step 3: The dose response tests based on the optimal contrasts
The estimates of the median time to vaginal delivery for each dose together with the optimal contrasts yield the adjusted p-values for testing the null hypothesis of no dose-response.
```{r mcttest, echo=FALSE}
f_mcttest <- function(dfobj){

  MCTtest <- MCTtest(doses,dfobj$est$median,
                     S=dfobj$est$S_median,
                     models=fmodels,
                     type="general",
                     contMat = dfobj$contmat$contmat,
                     pVal = TRUE, alternative="one.sided",
                     alpha=0.05,
                     critV=TRUE)
  p_table <- MCTtest$tStat %>%
    as_tibble(rownames=NA) %>%
    rownames_to_column(var="cm") %>%
    mutate(pval_ = attr(value,"pVal")) %>%
    mutate(pval = if_else(pval_<0.001, "<0.001",
                  if_else(pval_ <0.1, as.character(round(pval_,digits=3), digits=3),
                         as.character(round(pval_,digits=2),digits=2)))) %>%
    arrange(pval_) %>%
    select(-pval_) %>%
    knitr::kable(digits=2, align=c("lrr"),col.names = c("Candidate model", "t-statistic", "Adjusted p-value"))

  list(mcttest=MCTtest, p_table = p_table)

}

prim$mcttest <- f_mcttest(prim)
prim$mcttest$p_table


```

\newpage

## Step 4: Model selection

As pre-specified in the SAP, model selection will be based on Akaike's Information Criterion (AIC), with lower value indicating better fit.


```{r fitMod, echo=FALSE, message=FALSE, results=FALSE, include=FALSE}

f_fitmod <- function(dfobj){
  f <- function (model){
    fitmod <- fitMod(doses, dfobj$est$median, S=dfobj$est$S_median, model=model, type="general")
    plot <- plot(fitmod, CI=TRUE, plotData="meansCI",
                  ylab="Median time to vaginal delivery", xlab="Dose (mg/h)")

    aic <- gAIC(fitmod)
    ed50 <- ED(fitmod,0.5)
    ed70 <- ED(fitmod,0.7)
    ed90 <- ED(fitmod,0.9)
    fitparms <- c(aic, ed50, ed70, ed90)
    names(fitparms) <- c("AIC", "ED50", "ED70", "ED90")
    list(fitmod = fitmod, plot=plot, fitparms = fitparms)
  }

  fitmod <- lapply(fmodels_names,f)
  table <- tibble(name = names(fitmod$linear$fitparms),
                  linear = fitmod$linear$fitparms,
                  emax = fitmod$emax$fitparms,
                  sigEmax = fitmod$sigEmax$fitparms,
                  logistic = fitmod$logistic$fitparms) %>%
    gather(key=Model,value = value,-name) %>%
    spread(key=name, value=value) %>%
    arrange(AIC) %>%
    knitr::kable(col.names = c("Model", "AIC-value", "ED~50~","ED~70~","ED~90~"), digits = 2)

   list(fitmod = fitmod, table = table)
}


prim$fitmod <- f_fitmod(prim)



```

### Fitted model plots and model fit parameters

The dose response curves for each model are plotted below with 95% confidence intervals.


```{r , echo=FALSE, message=FALSE}
library(gridExtra)

grid.arrange(prim$fitmod$fitmod$linear$plot,
             prim$fitmod$fitmod$emax$plot,
             ncol=2)
grid.arrange(prim$fitmod$fitmod$sigEmax$plot,
             prim$fitmod$fitmod$logistic$plot,
             ncol=2)

```

Parameters showing model fit according to AIC with ED~50~, ED~70~ and ED~90~ is given by the table below.

```{r , echo=FALSE, message=FALSE}
prim$fitmod$table

```

## Target dose selection
The target dose will be based on estimates of ED~50~, ED~70~ and ED~90~ of the best fitted model together with the safety profile of each dose. The maximal dose (25 mg/h) will also be considered for target dose if there are no safety issues.


# Robustness analyses: Without covariates (FAS)

## Descriptives
Descriptives are the same as for the primary analysis

## Step 1: The candidate dose response models
Candidate dose response models are the same as for the primary analysis

## Step 2: Optimal contrasts from the candidate models
The transformed estimates of median time to delivery from the Weibull scale parameters and the variance-covariance matrix are given by

```{r , echo=FALSE, message=FALSE}

robust_nocovar <- list()

robust_nocovar$sreg <-survreg( Surv(aval, cnsr == 0)~ trtp -1, data = adprim, dist="weibull")
robust_nocovar$est <- f_est(adprim, robust_nocovar$sreg)

robust_nocovar$est$table %>%
  select(term,median_txt) %>%
  knitr::kable( col.names=c("Treatment" ,"Median time to delivery (95% CI)"), caption="Transformed estimates")

robust_nocovar$est$S_median %>%
  as_tibble() %>%
  rownames_to_column(var="S") %>%
  mutate(S = robust_nocovar$est$term) %>%
  rename_at(vars(V1:V4), funs(robust_nocovar$est$term)) %>%
  knitr::kable(digits=3, caption="Variance-covariance matrix")
```

Together with the candidate response models, the optimal contrasts are given by

```{r , echo=FALSE}
library(DoseFinding)

robust_nocovar$contmat <- f_contmat(robust_nocovar)
robust_nocovar$contmat$table
```

## Step 3: The dose response tests based on the optimal contrasts

The estimates of the median time to vaginal delivery for each dose together with the optimal contrasts yield the adjusted p-values for testing the null hypothesis of no dose-response.

```{r, echo=FALSE}

robust_nocovar$mcttest <- f_mcttest(robust_nocovar)
robust_nocovar$mcttest$p_table
```

## Step 4: Model selection

Model selection based on Akaike's Information Criterion (AIC), with lower value indicating better fit. The dose response curves for each model are plotted below with 95% confidence intervals.


```{r , echo=FALSE, message=FALSE, results=FALSE, include=FALSE}
robust_nocovar$fitmod <- f_fitmod(robust_nocovar)
```
```{r, echo=FALSE, message=FALSE}
gridExtra::grid.arrange(robust_nocovar$fitmod$fitmod$linear$plot,
             robust_nocovar$fitmod$fitmod$emax$plot,
             ncol=2)
gridExtra::grid.arrange(robust_nocovar$fitmod$fitmod$sigEmax$plot,
             robust_nocovar$fitmod$fitmod$logistic$plot,
             ncol=2)
```

Parameters showing model fit according to AIC with ED~50~, ED~70~ and ED~90~ is given by the table below.

```{r fitmodfit, echo=FALSE, message=FALSE}
robust_nocovar$fitmod$table

```

# Robustness analyses: Excluding women with caesarean sections (FAS)
This robustness analysis on the primary variable is performed on the women in FAS who did not deliver by caesarean section.



There were `r sum(adprim$fasfl=="Y" & adprim$cnsr != 1)` women included in the Full Analysis Set, and `r sum(adprim$pprotfl == "Y" & adprim$cnsr != 1)` women included in the Per Protocol Population who did not go to caesarean sections. 

```{r, echo = FALSE, warning = FALSE}
advds <- adprim %>%
  filter(cnsr != 1) # to be updated!!!
robust_vds <- list()
robust_vds$alloc <- f_alloc(advds)
robust_vds$alloc
```

## Descriptive statistics

```{r, echo=FALSE, message=FALSE}
robust_vds$desc <- f_km(advds)
robust_vds$desc$kmplot
robust_vds$desc$quantile

```

## Step 1: Candidate dose response models

Candidate dose response models are the same as for the primary analysis

## Step 2: Optimal contrasts from the candidate models

The transformed estimates of median time to delivery from the Weibull scale parameters and the variance-covariance matrix are given by

```{r , echo=FALSE, message=FALSE}


robust_vds$sreg <-survreg(Surv(aval, cnsr == 0)~ trtp -1 + indcdc + sitegrpc, data = advds, dist="weibull")
robust_vds$est <- f_est(advds, robust_vds$sreg)

robust_vds$est$table %>%
  select(term,median_txt) %>%
  knitr::kable( col.names=c("Treatment" ,"Median time to delivery (95% CI)"), caption="Transformed estimates")

robust_vds$est$S_median %>%
  as_tibble() %>%
  rownames_to_column(var="S") %>%
  mutate(S = robust_vds$est$term) %>%
  rename_at(vars(V1:V4), funs(robust_vds$est$term)) %>%
  knitr::kable(digits=3, caption="Variance-covariance matrix")
```

Together with the candidate response models, the optimal contrasts are given by

```{r , echo=FALSE}
library(DoseFinding)

robust_vds$contmat <- f_contmat(robust_vds)
robust_vds$contmat$table
```

## Step 3: The dose response tests based on the optimal contrasts

The estimates of the median time to vaginal delivery for each dose together with the optimal contrasts yield the adjusted p-values for testing the null hypothesis of no dose-response.

```{r, echo=FALSE}

robust_vds$mcttest <- f_mcttest(robust_vds)
robust_vds$mcttest$p_table
```

## Step 4: Model selection

Model selection based on Akaike's Information Criterion (AIC), with lower value indicating better fit. The dose response curves for each model are plotted below with 95% confidence intervals.


```{r , echo=FALSE, message=FALSE, results=FALSE, include=FALSE}
robust_vds$fitmod <- f_fitmod(robust_vds)
```
```{r, echo=FALSE, message=FALSE}
gridExtra::grid.arrange(robust_vds$fitmod$fitmod$linear$plot,
             robust_vds$fitmod$fitmod$emax$plot,
             ncol=2)
gridExtra::grid.arrange(robust_vds$fitmod$fitmod$sigEmax$plot,
             robust_vds$fitmod$fitmod$logistic$plot,
             ncol=2)
```

Parameters showing model fit according to AIC with ED~50~, ED~70~ and ED~90~ is given by the table below.

```{r , echo=FALSE, message=FALSE}
robust_vds$fitmod$table

```

# Robustness analyses: treating caesarean sections as non-censored events (FAS)

In this robustness analysis on the FAS, the infant's birth time for caesarean sections is not treated as censoring times, but as the primary event (vaginal delivery).

```{r, echo=FALSE, message = FALSE }
adcsnc <- adprim %>%
  mutate(cnsr = if_else(cnsr == 1, 0, cnsr))
robust_csnocens <- list()
```

## Descriptives

```{r, echo=FALSE, message=FALSE}
robust_csnocens$desc <- f_km(adcsnc)
robust_csnocens$desc$kmplot
robust_csnocens$desc$quantile

```

## Step 1: The candidate dose response models

Candidate dose response models are the same as for the primary analysis

## Step 2: Optimal contrasts from the candidate models

The transformed estimates of median time to delivery from the Weibull scale parameters and the variance-covariance matrix are given by

```{r , echo=FALSE, message=FALSE}

robust_csnocens <- list()

robust_csnocens$sreg <-survreg( Surv(aval, cnsr == 0)~ trtp -1 + indcdc + sitegrpc, data = adcsnc, dist="weibull")
robust_csnocens$est <- f_est(adcsnc, robust_csnocens$sreg)

robust_csnocens$est$table %>%
  select(term,median_txt) %>%
  knitr::kable( col.names=c("Treatment" ,"Median time to delivery (95% CI)"), caption="Transformed estimates")

robust_csnocens$est$S_median %>%
  as_tibble() %>%
  rownames_to_column(var="S") %>%
  mutate(S = robust_csnocens$est$term) %>%
  rename_at(vars(V1:V4), funs(robust_csnocens$est$term)) %>%
  knitr::kable(digits=3, caption="Variance-covariance matrix")
```

Together with the candidate response models, the optimal contrasts are given by

```{r , echo=FALSE}
library(DoseFinding)

robust_csnocens$contmat <- f_contmat(robust_csnocens)
robust_csnocens$contmat$table
```

## Step 3: The dose response tests based on the optimal contrasts

The estimates of the median time to vaginal delivery for each dose together with the optimal contrasts yield the adjusted p-values for testing the null hypothesis of no dose-response.

```{r, echo=FALSE}

robust_csnocens$mcttest <- f_mcttest(robust_csnocens)
robust_csnocens$mcttest$p_table
```

## Step 4: Model selection

Model selection based on Akaike's Information Criterion (AIC), with lower value indicating better fit. The dose response curves for each model are plotted below with 95% confidence intervals.


```{r , echo=FALSE, message=FALSE, results=FALSE, include=FALSE}
robust_csnocens$fitmod <- f_fitmod(robust_csnocens)
```
```{r, echo=FALSE, message=FALSE}
gridExtra::grid.arrange(robust_csnocens$fitmod$fitmod$linear$plot,
             robust_csnocens$fitmod$fitmod$emax$plot,
             ncol=2)
gridExtra::grid.arrange(robust_csnocens$fitmod$fitmod$sigEmax$plot,
             robust_csnocens$fitmod$fitmod$logistic$plot,
             ncol=2)
```

Parameters showing model fit according to AIC with ED~50~, ED~70~ and ED~90~ is given by the table below.

```{r , echo=FALSE, message=FALSE}
robust_csnocens$fitmod$table

```

# Primary analysis on the Per-Protocol analysis set

The Per-Protocol (PP) analysis set consists of women in the Full Analysis set who report no major protocol violations affecting the efficacy evaluation.


```{r, echo=FALSE, message = FALSE }
adprim_pp <- adprim %>%
  filter(pprotfl == "Y")
prim_pp <- list()
```

## Descriptives
Descriptive analysis of the primary variable on the PP analysis set includes Kaplan-Meier plots and corresponding quantiles by treatment group.

```{r, echo=FALSE, message=FALSE}
prim_pp$desc <- f_km(adprim_pp)
prim_pp$desc$kmplot
prim_pp$desc$quantile

```

## Step 1: The candidate dose response models

Candidate dose response models are the same as for the primary analysis

## Step 2: Optimal contrasts from the candidate models

The transformed estimates of median time to delivery from the Weibull scale parameters and the variance-covariance matrix are given by

```{r , echo=FALSE, message=FALSE}

prim_pp <- list()

prim_pp$sreg <-survreg( Surv(aval, cnsr == 0)~ trtp - 1 + indcdc + sitegrpc, data = adprim_pp, dist="weibull")
prim_pp$est <- f_est(adprim_pp, prim_pp$sreg)

prim_pp$est$table %>%
  select(term,median_txt) %>%
  knitr::kable( col.names=c("Treatment" ,"Median time to delivery (95% CI)"), caption="Transformed estimates")

prim_pp$est$S_median %>%
  as_tibble() %>%
  rownames_to_column(var="S") %>%
  mutate(S = prim_pp$est$term) %>%
  rename_at(vars(V1:V4), funs(prim_pp$est$term)) %>%
  knitr::kable(digits=3, caption="Variance-covariance matrix")
```

Together with the candidate response models, the optimal contrasts are given by

```{r , echo=FALSE}
library(DoseFinding)

prim_pp$contmat <- f_contmat(prim_pp)
prim_pp$contmat$table
```

## Step 3: The dose response tests based on the optimal contrasts

The estimates of the median time to vaginal delivery for each dose together with the optimal contrasts yield the adjusted p-values for testing the null hypothesis of no dose-response.

```{r, echo=FALSE}

prim_pp$mcttest <- f_mcttest(prim_pp)
prim_pp$mcttest$p_table
```

## Step 4: Model selection

Model selection based on Akaike's Information Criterion (AIC), with lower value indicating better fit. The dose response curves for each model are plotted below with 95% confidence intervals.


```{r , echo=FALSE, message=FALSE, results=FALSE, include=FALSE}
prim_pp$fitmod <- f_fitmod(prim_pp)
```
```{r, echo=FALSE, message=FALSE}
gridExtra::grid.arrange(prim_pp$fitmod$fitmod$linear$plot,
             prim_pp$fitmod$fitmod$emax$plot,
             ncol=2)
gridExtra::grid.arrange(prim_pp$fitmod$fitmod$sigEmax$plot,
             prim_pp$fitmod$fitmod$logistic$plot,
             ncol=2)
```

Parameters showing model fit according to AIC with ED~50~, ED~70~ and ED~90~ is given by the table below.

```{r , echo=FALSE, message=FALSE}
prim_pp$fitmod$table

```

# Robustness analysis on the Per-Protocol analysis set with women with caesarean sections excluded

This robustness analysis on the primary variable is perfomed on the women in the Per-Protocol analysis set who did not deliver by caesarean section.


```{r, echo=FALSE, message = FALSE }
advds_pp <- adprim %>%
  filter(pprotfl == "Y" & cnsr != 1) # to be updated!!!
robust_vds_pp <- list()
```

## Descriptives
Descriptive analysis of the primary variable on the PP analysis set includes Kaplan-Meier plots and corresponding quantiles by treatment group.

```{r, echo=FALSE, message=FALSE}
robust_vds_pp$desc <- f_km(advds_pp)
robust_vds_pp$desc$kmplot
robust_vds_pp$desc$quantile

```

## Step 1: The candidate dose response models

Candidate dose response models are the same as for the primary analysis

## Step 2: Optimal contrasts from the candidate models

The transformed estimates of median time to delivery from the Weibull scale parameters and the variance-covariance matrix are given by

```{r , echo=FALSE, message=FALSE}

robust_vds_pp <- list()

robust_vds_pp$sreg <-survreg( Surv(aval, cnsr == 0)~ trtp - 1 + indcdc + sitegrpc, data = advds_pp, dist="weibull")
robust_vds_pp$est <- f_est(advds_pp, robust_vds_pp$sreg)

robust_vds_pp$est$table %>%
  select(term,median_txt) %>%
  knitr::kable( col.names=c("Treatment" ,"Median time to delivery (95% CI)"), caption="Transformed estimates")

robust_vds_pp$est$S_median %>%
  as_tibble() %>%
  rownames_to_column(var="S") %>%
  mutate(S = robust_vds_pp$est$term) %>%
  rename_at(vars(V1:V4), funs(robust_vds_pp$est$term)) %>%
  knitr::kable(digits=3, caption="Variance-covariance matrix")
```

Together with the candidate response models, the optimal contrasts are given by

```{r , echo=FALSE}
library(DoseFinding)

robust_vds_pp$contmat <- f_contmat(robust_vds_pp)
robust_vds_pp$contmat$table
```

## Step 3: The dose response tests based on the optimal contrasts

The estimates of the median time to vaginal delivery for each dose together with the optimal contrasts yield the adjusted p-values for testing the null hypothesis of no dose-response.

```{r, echo=FALSE}

robust_vds_pp$mcttest <- f_mcttest(robust_vds_pp)
robust_vds_pp$mcttest$p_table
```

## Step 4: Model selection

Model selection based on Akaike's Information Criterion (AIC), with lower value indicating better fit. The dose response curves for each model are plotted below with 95% confidence intervals.


```{r , echo=FALSE, message=FALSE, results=FALSE, include=FALSE}
robust_vds_pp$fitmod <- f_fitmod(robust_vds_pp)
```
```{r, echo=FALSE, message=FALSE}
gridExtra::grid.arrange(robust_vds_pp$fitmod$fitmod$linear$plot,
             robust_vds_pp$fitmod$fitmod$emax$plot,
             ncol=2)
gridExtra::grid.arrange(robust_vds_pp$fitmod$fitmod$sigEmax$plot,
             robust_vds_pp$fitmod$fitmod$logistic$plot,
             ncol=2)
```

Parameters showing model fit according to AIC with ED~50~, ED~70~ and ED~90~ is given by the table below.

```{r , echo=FALSE, message=FALSE}
robust_vds_pp$fitmod$table

```


 
