---
title: "ImPRESS — Time-to-Event Proportional-Hazards Assessment"
author: "ImPRESS Statistical Team"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: xelatex
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 7, fig.height = 4.5, out.width = "100%")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

library(targets)
library(dplyr)
library(survival)
library(survminer)
library(broom)
library(tibble)
library(knitr)

# Strip element_markdown themes that survminer sometimes attaches (they do not
# render under the standard graphics device). Mirrors make_rdtte.R.
sanitize_theme <- function(p) {
  if (is.null(p) || is.null(p$theme)) return(p)
  th <- p$theme
  for (nm in names(th)) {
    if (inherits(th[[nm]], "element_markdown")) {
      th[[nm]] <- ggplot2::element_text()
    }
  }
  p$theme <- th
  p
}
```

```{r load-data}
tar_load(adtte)
tar_load(cfg)

cohortcd_analysis <- as.character(cfg$analysis$cohortcd)
adtte_cohort <- adtte %>%
  dplyr::filter(cohortcd == cohortcd_analysis)

endpoints <- adtte_cohort %>%
  dplyr::distinct(paramcd, param) %>%
  dplyr::arrange(paramcd)
n_subj <- dplyr::n_distinct(adtte_cohort$usubjid)
```

# Introduction

This document provides a pre-unblinding assessment of the **proportional-hazards (PH)
assumption** for the Cox models reported in the time-to-event section of the main
mock-up report (`make_rdtte.R`), in **`r n_subj` subjects** (cohort
`r cohortcd_analysis`). It addresses consistency-review point **C3** (SAP states the
PH assumption is assessed by standard diagnostics; the main report shows hazard ratios
without the diagnostics).

For each endpoint the Cox model is `coxph(Surv(aval, 1 - cnsr) ~ trtcd)`, where `trtcd`
is the long-term planned dose (0/25/50/100 mg) — the same specification used for the
reported hazard ratios. The PH assumption is assessed with:

1. **Scaled Schoenfeld residual test** (`survival::cox.zph`) — a global test plus a
   per-term test; small p-values indicate a time-varying (non-proportional) effect.
2. **Schoenfeld residual plots** — scaled residuals against transformed time; a
   non-zero slope signals departure from PH.
3. **Complementary log-log survival plots** — `log(-log S(t))` against `log(time)` by
   dose; approximately parallel curves are consistent with PH.

All analyses use sham-randomised data and are descriptive.

> **Interpretation guide.** Treat the global Schoenfeld p-value as the headline:
> p > 0.10 — no evidence against PH; 0.05–0.10 — borderline, inspect plots;
> p < 0.05 — evidence of non-proportionality, consider a time-varying coefficient,
> stratification, or reporting a time-averaged hazard ratio with this caveat.

---

# Per-endpoint diagnostics

```{r ph-diagnostics, results='asis', fig.height=5}
zph_summary <- list()

for (i in seq_len(nrow(endpoints))) {
  pcd <- endpoints$paramcd[i]
  lbl <- endpoints$param[i]

  dt <- adtte_cohort %>%
    dplyr::filter(paramcd == pcd, !is.na(aval), !is.na(cnsr)) %>%
    dplyr::mutate(
      trtcd = factor(dose02p, levels = c(0, 25, 50, 100)),
      trt   = factor(paste0(trtcd, " mg"), levels = c("0 mg", "25 mg", "50 mg", "100 mg"))
    ) %>%
    dplyr::filter(!is.na(trtcd))

  cat("\n\n## ", lbl, " (`", pcd, "`)\n\n", sep = "")

  n_events <- sum(dt$cnsr == 0L, na.rm = TRUE)
  if (nrow(dt) == 0 || dplyr::n_distinct(dt$trtcd) < 2 || n_events < 5) {
    cat("Insufficient data for a PH assessment (",
        n_events, " events, ", dplyr::n_distinct(dt$trtcd),
        " dose groups).\n\n", sep = "")
    next
  }

  cat(n_events, " events in ", nrow(dt), " subjects.\n\n", sep = "")

  cox <- tryCatch(
    survival::coxph(survival::Surv(aval, 1L - cnsr) ~ trtcd, data = dt),
    error = function(e) NULL
  )
  if (is.null(cox)) { cat("Cox model failed to fit.\n\n"); next }

  zph <- tryCatch(survival::cox.zph(cox), error = function(e) NULL)
  if (is.null(zph)) { cat("Schoenfeld test could not be computed.\n\n"); next }

  # ---- Schoenfeld test table ----------------------------------------------
  zph_tbl <- as.data.frame(zph$table)
  zph_tbl$Term <- rownames(zph_tbl)
  zph_tbl <- zph_tbl %>%
    dplyr::transmute(
      Term = dplyr::recode(Term, GLOBAL = "GLOBAL (overall)"),
      chisq = round(chisq, 3),
      df = df,
      p = round(p, 4)
    )
  global_p <- zph$table["GLOBAL", "p"]
  zph_summary[[pcd]] <- tibble::tibble(
    paramcd = pcd, param = lbl, global_p = round(global_p, 4),
    conclusion = dplyr::case_when(
      is.na(global_p)     ~ "Not estimable",
      global_p > 0.10     ~ "No evidence against PH",
      global_p > 0.05     ~ "Borderline",
      TRUE                ~ "Evidence of non-proportionality"
    )
  )

  print(knitr::kable(zph_tbl,
                     col.names = c("Term", "Chi-square", "df", "p-value"),
                     caption = paste0("Scaled Schoenfeld test — ", lbl)))
  cat("\n\n")

  cat("**Global Schoenfeld p-value:** ",
      ifelse(is.na(global_p), "NA",
             ifelse(global_p < 0.001, "<0.001", format(round(global_p, 3), nsmall = 3))),
      " — ", zph_summary[[pcd]]$conclusion, ".\n\n", sep = "")

  # ---- Schoenfeld residual plot -------------------------------------------
  cat("### Scaled Schoenfeld residuals\n\n")
  sp <- tryCatch(survminer::ggcoxzph(zph), error = function(e) NULL)
  if (!is.null(sp)) {
    print(sp)
  } else {
    op <- graphics::par(no.readonly = TRUE)
    plot(zph)
    graphics::par(op)
  }
  cat("\n\n")

  # ---- Complementary log-log plot -----------------------------------------
  cat("### Complementary log-log survival\n\n")
  fit <- survival::survfit(survival::Surv(aval, 1L - cnsr) ~ trt, data = dt)
  cl <- tryCatch(
    survminer::ggsurvplot(
      fit, data = dt, fun = "cloglog",
      legend.title = "Dose", legend.labs = levels(droplevels(dt$trt)),
      xlab = "log(days since randomisation)", ylab = "log(-log S(t))"
    ),
    error = function(e) NULL
  )
  if (!is.null(cl)) {
    cl$plot <- sanitize_theme(cl$plot)
    print(cl$plot)
  }
  cat("\n\n---\n\n")
}
```

# Summary

```{r ph-summary}
if (length(zph_summary) == 0) {
  cat("No endpoints had sufficient data for a PH assessment.\n")
} else {
  dplyr::bind_rows(zph_summary) %>%
    knitr::kable(
      col.names = c("PARAMCD", "Parameter", "Global Schoenfeld p", "Conclusion"),
      caption = "Global proportional-hazards test by endpoint"
    )
}
```

Endpoints flagged as borderline or non-proportional should be interpreted with the
corresponding plots; for those a time-averaged hazard ratio is reported in the main
document with this caveat, and a time-varying-coefficient or stratified sensitivity
analysis can be added at database lock.

---

*Generated from `reports/model fit/impress_tte_ph_check.Rmd` — sham-randomised data.*
